---
title: "EDA"
format: html
---

```{r message = FALSE}
library(tidyverse)
library(ggplot2)
library(dplyr)
```

DATA:

NACCAGE

EDUC

SEX

NACCUDSD

```{r}
mri_cleaned <- read.csv("C:\\Alzheimer Project\\original_data\\mri.csv")
uds_cleaned <- read.csv("C:\\Alzheimer Project\\original_data\\uds.csv")
mri_raw <- read.csv("C:\\Alzheimer Project\\original_data\\investigator_fcsf_nacc57.csv")
uds_raw <- read.csv("C:\\Alzheimer Project\\original_data\\investigator_uds_baseline.csv")
```


```{r}
data_cleaned <- uds_cleaned %>% inner_join(mri_cleaned, by = "NACCID")
data_raw <- uds_raw %>% inner_join(mri_raw, by = "NACCID")
```


Distribution of Age: 

```{r}
ggplot(data=uds_cleaned, aes(x = NACCAGE)) +
  geom_histogram(col = I("white"))
```


Distribution of Education:

```{r}
ggplot(data=uds_cleaned, aes(x = EDUC)) +
  geom_histogram(col = I("white"))
```

Distribution of Sex:

```{r}
uds_cleaned %>% filter(SEX==1) %>% nrow()
uds_cleaned %>% filter(SEX==2) %>% nrow()
```


```{r}
ggplot(data=uds_cleaned, aes(x = "", fill = factor(SEX))) + 
  geom_bar(width = 1) +
  coord_polar(theta = "y") +
  labs(title = "Distribution of Gender", 
       fill = "Sex") +
  scale_fill_discrete(labels = c("Male", "Female"))
```

1 = Male

2 = Female

Distribution of Cognitive Status at UDS visit:

```{r}
ggplot(data=uds_cleaned, aes(x = factor(NACCUDSD))) + 
  geom_bar() +
  labs(x = "Cognitive Status at UDS visit") +
  theme_bw()
```

1 = Normal cognition 

2 = Impaired-not-MCI 

3 = MCI 

4 = Dementia

The association between age and sex:

```{r}
ggplot(data=uds_cleaned, aes(x = NACCAGE, y = factor(SEX))) +
  geom_boxplot() +
  labs(y = "Sex")
```

1 = Male

2 = Female

The association between age and education:

```{r}
ggplot(data=uds_cleaned, aes(x = NACCAGE, y = EDUC)) +
  geom_point()
```

The association between sex and education: 

```{r}
ggplot(data=uds_cleaned, aes(x = factor(SEX), y = NACCAGE)) + 
  geom_boxplot() +
  labs(x = "Sex")
```

1 = Male

2 = Female

The relationship between age and cognitive status:

```{r}
ggplot(data=uds_cleaned, aes(x=NACCAGE, y =factor(NACCUDSD))) +
  geom_boxplot() +
  labs(y = "Cognitive Status at UDS visit")
```

1 = Normal cognition 

2 = Impaired-not-MCI 

3 = MCI 

4 = Dementia

Sex and cognitive status:

```{r}
ggplot(data=uds_cleaned, aes(x = factor(NACCUDSD), fill = factor(SEX))) +
  geom_bar(position= "dodge") +
  labs(x = "Cognitive Status at UDS visit", fill = "Sex") +
  scale_x_discrete(labels = c("HC", "Impaired-not-MCI", "MCI", "AD")) +
  scale_fill_discrete(labels = c("Male", "Female"))
```

1 = Normal cognition 

2 = Impaired-not-MCI 

3 = MCI 

4 = Dementia

Education and cognitive status:

```{r}
ggplot(data=uds_cleaned, aes(x = EDUC, y = factor(NACCUDSD))) + 
  geom_boxplot() +
  labs(x = "Years of Education", y = "Cognitive Status at UDS visit")
```

1 = Normal cognition 

2 = Impaired-not-MCI 

3 = MCI 

4 = Dementia


Age, Sex, and Cognitive Status: 

```{r}
ggplot(data=uds_cleaned, aes(x=NACCAGE, y =factor(NACCUDSD))) +
  geom_boxplot() +
  facet_grid(SEX~.) +
  labs(y = "Cognitive Status at UDS visit") 
```


BEHAVIORAL SURVEY DATA: (cleaned data) 

TRAVEL

EVENTS 

REMDATES

MEALPREP

SHOPPING

BILLS

PAYATTN

TAXES

STOVE



```{r}
ggplot(data=uds_cleaned, aes(x = factor(NACCUDSD), fill = factor(TRAVEL))) +
  geom_bar(position = "dodge") +
  labs(x = "Cognitive Status",
       fill = "Travel Ability") +
  scale_x_discrete(labels = c("HC", "Impaired-not-MCI", "MCI", "AD")) +
  scale_fill_discrete(labels = c("Normal", "Has difficulty, but does by self","Requires assistance", "Dependent")) +
  theme_bw()
```

1 = Normal cognition 

2 = Impaired-not-MCI 

3 = MCI 

4 = Dementia

```{r}
ggplot(data=uds_cleaned, aes(x = factor(NACCUDSD), 
                             fill = factor(EVENTS))) +
  geom_bar(position = "dodge") +
  labs(x = "Cognitive Status",
       fill = "Ability to keep track of current events") +
  scale_x_discrete(labels = c("HC", "Impaired-not-MCI", "MCI", "AD")) +
  scale_fill_discrete(labels = c("Normal", "Has difficulty, but does by self","Requires assistance", "Dependent")) +
  theme_bw()
```

1 = Normal cognition 

2 = Impaired-not-MCI 

3 = MCI 

4 = Dementia

```{r}
ggplot(data=uds_cleaned, aes(x = factor(NACCUDSD), fill = factor(REMDATES))) +
  geom_bar(position = "dodge") +
  labs(x = "Cognitive Status",
       fill = "Ability to remember dates") +
  scale_x_discrete(labels = c("HC", "Impaired-not-MCI", "MCI", "AD")) +
  scale_fill_discrete(labels = c("Normal", "Has difficulty, but does by self","Requires assistance", "Dependent")) +
  theme_bw()
```

1 = Normal cognition 

2 = Impaired-not-MCI 

3 = MCI 

4 = Dementia

```{r}
ggplot(data=uds_cleaned, aes(x = factor(NACCUDSD), fill = factor(MEALPREP))) +
  geom_bar(position = "dodge") +
  labs(x = "Cognitive Status",
       fill = "Ability to prepare balanced meals") +
  scale_x_discrete(labels = c("HC", "Impaired-not-MCI", "MCI", "AD")) +
  scale_fill_discrete(labels = c("Normal", "Has difficulty, but does by self","Requires assistance", "Dependent")) +
  theme_bw()
```

1 = Normal cognition 

2 = Impaired-not-MCI 

3 = MCI 

4 = Dementia


```{r}
ggplot(data=uds_cleaned, aes(x = factor(NACCUDSD), fill = factor(SHOPPING))) +
  geom_bar(position = "dodge") +
  labs(x = "Cognitive Status",
       fill = "Ability to buy groceries/go shopping") +
  scale_x_discrete(labels = c("HC", "Impaired-not-MCI", "MCI", "AD")) +
  scale_fill_discrete(labels = c("Normal", "Has difficulty, but does by self","Requires assistance", "Dependent")) +
  theme_bw()
```

1 = Normal cognition 

2 = Impaired-not-MCI 

3 = MCI 

4 = Dementia


```{r}
ggplot(data=uds_cleaned, aes(x = factor(NACCUDSD), 
                             fill = factor(PAYATTN))) +
  geom_bar(position = "dodge") +
  labs(x = "Cognitive Status",
       fill = "Ability to pay atttention") +
  scale_x_discrete(labels = c("HC", "Impaired-not-MCI", "MCI", "AD")) +
  scale_fill_discrete(labels = c("Normal", "Has difficulty, but does by self","Requires assistance", "Dependent")) +
  theme_bw()
```

1 = Normal cognition 

2 = Impaired-not-MCI 

3 = MCI 

4 = Dementia

```{r}
ggplot(data=uds_cleaned, aes(x = factor(NACCUDSD), 
                             fill = factor(BILLS))) +
  geom_bar(position = "dodge") +
  labs(x = "Cognitive Status",
       fill = "Ability to write checks/pay bills") +
  scale_x_discrete(labels = c("HC", "Impaired-not-MCI", "MCI", "AD")) +
  scale_fill_discrete(labels = c("Normal", "Has difficulty, but does by self","Requires assistance", "Dependent"))+
  theme_bw()
```

```{r}
ggplot(data=uds_cleaned, aes(x = factor(NACCUDSD), 
                             fill = factor(TAXES))) +
  geom_bar(position = "dodge") +
  labs(x = "Cognitive Status",
       fill = "Ability to do tax documents") +
  scale_x_discrete(labels = c("HC", "Impaired-not-MCI", "MCI", "AD")) +
  scale_fill_discrete(labels = c("Normal", "Has difficulty, but does by self","Requires assistance", "Dependent")) +
  theme_bw()
```

1 = Normal cognition 

2 = Impaired-not-MCI 

3 = MCI 

4 = Dementia


```{r}
ggplot(data=uds_cleaned, aes(x = factor(NACCUDSD), 
                             fill = factor(STOVE))) +
  geom_bar(position = "dodge") +
  labs(x = "Cognitive Status",
       fill = "Ability to heat water/
       make coffee/
       turn off the stove") +
  scale_x_discrete(labels = c("HC", "Impaired-not-MCI", "MCI", "AD")) +
  scale_fill_discrete(labels = c("Normal", "Has difficulty, but does by self","Requires assistance", "Dependent")) +
  theme_bw()
```

1 = Normal cognition 

2 = Impaired-not-MCI 

3 = MCI 

4 = Dementia


```{r}
ggplot(data=uds_cleaned, aes(x = factor(NACCUDSD), 
                             fill = factor(GAMES))) +
  geom_bar(position = "dodge") +
  labs(x = "Cognitive Status",
       fill = "Ability to play games/work on a hobby") +
  scale_x_discrete(labels = c("HC", "Impaired-not-MCI", "MCI", "AD")) +
  scale_fill_discrete(labels = c("Normal", "Has difficulty, but does by self","Requires assistance", "Dependent")) +
  theme_bw()
```

1 = Normal cognition 

2 = Impaired-not-MCI 

3 = MCI 

4 = Dementia



